---
title: "3) Different Guidance Functioning"
author: "Florian Berding and Julia Pargmann"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{03_dgf}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1 Introduction

In this vignette, we would like to show you how the tools of the *Iota
Concept* can be used to analyze the suitability of a coding scheme for
rating different kinds of materials. The analysis is very similar to the
analysis of different item functioning (DIF) from item-response theory.
The focal point here is the question whether or not a coding scheme has
the same degree of reliability for different groups of materials. This
issue is closely connected to the question if a coding scheme reproduces
bias for specific groups.

If a coding scheme is perceived as a measurement instrument, like a test
or questionnaire, it must be reliable and valid. Reliability and
validity demand the absence of different guidance functioning (DGF), since
the existence of DGF means that the data is influenced by other sources, as
for example the phenomenon of interest.

We use the term "different guidance functioning" in order to eliminate
this kind of analysis from DIF. Furthermore, we would like to express
that a coding scheme usually only
provides a guideline for raters in assessing a phenomenon.

The analysis of DIF is important for a broad range of applications. For
example,

-   ensuring that ratings of newspapers do not prefer or discourage
    specific political parties.
-   ensuring that books in different languages are rated without the
    influence of the different languages.
-   ensuring that social media contributions are rated similarly without
    bias for the platform-specific form of contribution.
-   ensuring that material created by different groups of learners is
    assessed without bias for their socioeconomic background.
-   ...

In the following, we would like to illustrate this kind of analysis by
continuing the example from the first [vignette](iotarelr.html). In
order to conduct an analysis of DGF, we need additional data that
contains information on how the different coding units are grouped. Let
us have a look at the sample data set.

# 2 Example Analysis
## 2.1 Basic Analysis

```{r}
library(iotarelr)
head(iotarelr_written_exams)
```

The data set contains the ratings of exams from three raters and an
additional column called "Sex". This column stores the information if
the participant of the exam identifies as male or a female. Since an
exam should only measure participants' performance and should be fair,
the *Assignment Error Matrix* in both groups should not differ. That is,
men and women should have the same chance for a good, an average or a
poor exam.

The corresponding analysis can be requested via the function
`check_dgf()`. This function is similar to `compute_iota2()` but has an
additional argument, `splitcr`. This argument uses the group information
for the analysis while `data` only uses the ratings. Please note that
calling this function may take some time.

```{r}
dgf_exam<-check_dgf(data=iotarelr_written_exams[c("Coder A","Coder B","Coder C")],
                    splitcr = iotarelr_written_exams$Sex,
                    random_starts = 300)
```

The results are stored in the object `dgf_exam`. For every group
specified in the column "Sex", a model of Iota2 is fitted. The results
can be accessed in the same way as with objects created with
`compute_iota2()`. The only thing to do is to request the function
`get_summary()`with the results for the specific group. Let us start with
the females.

```{r}
get_summary(dgf_exam$group_female)
```

The summary for the females shows the basic information of the fitted
model. The best log-likelihood has been replicated. Thus, we can be
confident that our model represents the best model for the females.

Special attention should be paid to the *Assignment Error Matrix*, since
this matrix describes how the true categories are assigned. For the
females, an average exam is recognized as indeed an average exam in about
88.7% of cases. A good exam is recognized as  good exam in about 50% of
cases. The remaining 50% of cases are assigned to be an average exam or a
a poor exam with the same probability of 25%. A truly poor exam is
assigned to be a poor exam in about 50% of cases and an average exam in nearly all other cases. The error to assign a poor exam to be a good exam
is only about 5%. Let us now have a look at the results for the men.

```{r}
get_summary(dgf_exam$group_male)
```

For the men, the result is very different. According to their
*Assignment Error Matrix*, an average exam is recognized as an average
exam only in about 50% of the cases. If it is not recognized as an average
exam, it has a higher chance to be considered a good exam than a
poor one. Concentrating on the good exams, the *Assignment Error Matrix*
shows the same probability for all categories. This implies that a good
exam is assigned randomly to any of the three categories. Only the poor
exams have a high chance to be assigned to the correct category with
more than 70%.

Comparing the *Assignment Error Matrices*, it becomes clear that the
coding scheme is more reliable for females than for men.

How the different degrees of reliability affect the labeled data can be
explored with the function `plot_iota()`. Here we can directly pass the
object `dgf_exam` to the function.

```{r, fig.height = 6, fig.width = 7.2, fig.align = "center", fig.cap = "Figure 1: Plot of Iota for different Groups"}
plot_iota(dgf_exam,
          ylab = "Groups")
```

Let us start with the average exams of the females. The red rectangle
for this category implies that the data labeled as "average" contains a
lot of exams which truly are either good or poor exams (*Iota Error
II*). The red rectangle is even bigger than the green (*Iota*) and
orange ones (*Iota Error I*), meaning that the number of averagely good
exams is overestimated for the females. Concentrating on the males, the
red rectangle (*Iota Error II*) is quite small but the orange one (*Iota
Error I*) is quite big. Thus, the number of averagely good exams is
underestimated.

Regarding the good exams of the females, the red rectangle is quite
small (*Iota Error II*). This means that there is only a small number of
average or poor exams that are labeled as good. However, the orange
rectangle is quite big (*Iota Error I*). It nearly has the same size as
the green rectangle (*Iota*). Thus, the data for the good exams misses
about half of the corresponding exams. This means that the number of
good exams is underestimated in the data set for the females.
Concentrating on the males, *Iota* is only about .18, which is low.
Thus, the good exams are not reliably represented in the individual
data. In contrast, about 36% of the labeled data are missing good exams
(*Iota Error I*). Instead, the data set is made up of about 46% of exams
from other categories (*Iota Error II*). That is, from truly average or
truly poor exams. These exams compensate the missing good exams. As a
consequence, in the *average of all male participants*, the number of
good exams is actually quite correct for the males. On the *level of
single males*, the data is not reliable.

Finally, both diagrams for the poor exams look similar for both men and
women. For the women, the number of missing poor exams is higher than
for the men (*Iota Error I*). In contrast, the data labeled as poor
exams contains more exams which truly belong to the category "good" or
"average" for the men as well as for the females (*Iota Error II*). Thus, in
*average over all female participants*, the number of poor exams is
quite correct as well. For the males, the number of poor exams is
overestimated. For men as well as for women, the data is not reliable on
the individual level.

## 2.2 Coding Stream Analysis
Since version 0.1.3, *iotarelr* provides an additional analysis of the 
coding stream that is able to make the interpretation of the data more
convenient. This kind of analysis can be request via the function `plot_iota2_alluvial()`. 
Please note that the estimates cannot be passed
directly to the function. Only the results for a single group are compatible.
Applied to the current example, Figure 2 shows the coding stream analysis for the
females. The interpretation of the plot is explained in the vignette 
[Get started](iotarelr.html).

```{r, fig.height = 5, fig.width = 7.2, fig.align = "center", fig.cap="Figure 2: Stream Plot for Females"}
plot_alluvial_females<-plot_iota2_alluvial(
  dgf_exam$group_female,
  label_titel = "Coding Stream from True to Assigned Categories (Females)")
plot(plot_alluvial_females)
```
Comparing the boxes for the true and the labeled data, it becomes clear that 
the number of average exams is overestimated while the number of good exams
is underestimated. The sizes of the boxes for the true and assigned poor exams
are similar. That is, on average, the number of poor exams is represented in the
labeled date quite correctly.

The stream plot illustrates that the exams labeled as good exams are made up by exams which are in truth good exams to a very high extent. This can be seen by the
big green curve and the blue and red curves, which in comparison are small. The exams labeled as poor exams are made up of poor and good exams to about 50%. The data labeled as average exams is made up by exams which are in truth good,
average and poor to about one third.

```{r, fig.height = 5, fig.width = 7.2, fig.align = "center", fig.cap="Figure 3: Stream Plot for Males"}
plot_alluvial_males<-plot_iota2_alluvial(
  dgf_exam$group_male,
  label_titel = "Coding Stream from True to Assigned Categories (Males)")
plot(plot_alluvial_males)
```

Turning to the males, the stream plot is very different. Comparing the boxes for
the true and assigned categories, it becomes clear that the number of average 
exams is underestimated while the poor exams are overestimated. Furthermore, 
the coding scheme leads to a small overestimation of the good exams. Thus,
the coding stream plot reflects the results from section 2.1, indicating that
the coding scheme tends to divide males' exams into good or poor exams with no
average in between, compared to an error-free measurement. 

# 3 Summary

To sum up, the coding scheme treats the exams of males and females
differently. The number of averagely good exams is overestimated for
females and underestimated for men. The number of good exams is
underestimated for females but quite accurate for men, although the
number has its source in the wrong exam categories. Finally, the number
of poor exams is quite correct for females, although it sources from
quite a lot of wrong exams, while the number of poor exams for the men
tends to be overestimated. Thus, in descriptive statistics, females
appear to have lower performance than they truly have. For men, the data
shows the tendency that they show either a good or a poor performance
but nothing in between.

This example shows how the guidance of the coding scheme can be analyzed
for different groups of materials. This information can help in
developing coding schemes that are free from bias, providing a more
reliable and valid source of data for subsequent analysis and
conclusions.
