---
title: "3) Different Guidance Functioning"
author: "Florian Berding and Julia Pargmann"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dgf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# 1 Introduction
In this vignette we would like to show you how the tools of the Iota Concept can
be used for analyzing the guidance of a coding scheme for rating different
kind of materials. The analysis is very similar to the analysis of Different
Item Functioning (DIF) from Item Response Theory. The focal point here is the
question if a coding scheme has the same degree of reliability for different
groups of material. This issue is closely connected to the question if
a coding scheme produces bias for specific groups. 

If a coding scheme is perceived as a measurement instrument like a test or questionnaire
it mus be reliable and valid. Reliability and validity demand the absence of
Different Guidance Function (DGF) since the existence of DGF means that data
is influenced by other sources as the phenomenon of interest. 

We use the term Different Guidance Functioning in order to delimitate this kind
of analysis from Different Item Functioning (DIF) which is part of the Item
Response Theory. Furthermore, we would like to express that a coding scheme 
provides a guideline for raters in assessing a phenomenon.

Analysis of Different Guidance Functioning is important for a broad range of
applications. For example,

- ensuring that ratings of newspapers do not prefer or discourage specific 
political parties.
- ensuring that books in different languages are rated without the influence
of the different languages.
- ensuring that social media contributions are rated similar without bias
from the platform specific form of the contribution.
- ensuring that material created by different groups of learners is assessed
without bias from their social-economic background.
- ...

In the following we would like to illustrate this kind of analysis by continuing
the example from the first [vignette](iotarelr.html). In order to conduct an
analysis of Different Guidance Functioning we need additional data that contains
the information how the different codings units are grouped. Let's have a look
into  current data set.

# 2 Example Analysis
```{r}
library(iotarelr)
head(iotarelr_written_exams)
```
The data set contains the ratings of exams from three raters and an additional 
column called "Gender". This column stores the information if the creator of
an exam was a men or a women. Since an exam should only measure participants'
performance and should be fair the *Assignment Error Matrix* in both groups
should not differ. That is men and women should have the same chance for a good,
an average or a poor exam.

The corresponding analysis can be requested via the function `check_dgf()` This
function is similar to `compute_iota2()` but has an additional argument `splitcr`.
This argument takes the group information for the analysis while `data` takes
only the ratings. Please note that calling this function may take some time. 

```{r}
dgf_exam<-check_dgf(data=iotarelr_written_exams[c("Coder A","Coder B","Coder C")],
                    splitcr = iotarelr_written_exams$Gender,
                     random_starts = 2)
```
The results are stored in the object *dgf_exam*. For every group specified in
the column "Gender" a model of Iota2 is fitted. The results an be accessed in the
same way as for objects created with `compute_iota2()`. The only thing to do is to
request the function `get_summary()`with the results for the specific group. Let's
start with the females.
```{r}
get_summary(dgf_exam$group_female)
```
The summary for the females show the basic information of the fitted model. The
best log likelihood has been replicated. Thus, we can be confident that our model
represents the best model for the females.

Special attention should be paid to the *Assignment Error Matrix* since this matrix
describes how the true categories are assigned. For the females an average exam is
assigned as an average exam in about 88.7 % of the cases. A good exam is assigned
as a good exam in about 50 % of the cases. The remaining 50 % of cases are assigned as
an average exam or as an poor exam with the same probability of 25 %. A truly poor
exam is assigned as a poor exam in about 50 % of the cases and nearly in all other
cases as an average exam. The error to assign a poor exam as a good exam is only 
about 5 %. Let's now have a look on the results for the men.
```{r}
get_summary(dgf_exam$group_male)
```
For the men the result is very different. According to their *Assignment Error Matrix*
 an average exam is assigned as an average exam only in 54.4 % of the cases.
If it is not recognized as an average exam it has a higher chance to be considered
as a good exams than as a poor exam. Concentrating on the good exams the
 *Assignment Error Matrix* shows the same probability for all categories. This implies that a good exam
is assigned randomly to any of the three categories. Only the poor exams have a high
chance to be assigned to the correct category with about 75.1 %. That is, a poor
exam is assigned as a poor exam in three from four cases. 

Comparing the *Assignment Error Matrices* it becomes 
clear that the coding scheme is more reliable for the females than for the men.

How the different degrees of reliability effect the generated data can be explored
with the function `plot_iota()`.
```{r, fig.height = 3, fig.width = 7.2, fig.align = "center"}
plot_iota(dgf_exam)
```
Let`s start with the average exams of the females. The red rectangle for this 
category implies that the data labeled as "average" contains a lot of exams which 
truly are good or poor exams. The red rectangle is even greater as the green 
and orange rectangle meaning that the number of average exams is overestimated
for the females. Concentrating on the males the red rectangle is quite small but
the organ rectangle is quite high. Thus, the number of average exams is underestimated.

Regarding the good exams of the females the red rectangle is quite small This means
that there is only a small number of average or poor exams which are labeled as good.
However, the orange rectangle is quite high. It has nearly the same size as the green
rectangle. Thus, the data for the good exams misses about the half of the corresponding
exams. Thus, the number of good exams is underestimated in the data set for the females.
Concentrating on the males *Iota* is only about .189 which is low. Thus, the good exams
are not reliable represented in the individual data. In contrast 37.7 %
of the relevant good exams are not recognized. Instead, the data set is made up to
43.4 % by exams from other categories. That is, from truly average or  truly poor exams.
These exams compensate the missing good exams. As a consequence, in the 
*average of all male participants* the number of good exams is quite correct for the males. On
the *level of single males* the data is not reliable. 

Finally, both graphics for the poor exams look similar for the men and women. For
the women, the number of missing true exams is higher as for the men. In contrast,
the data labeled as poor exams contains more exams which truly belong to the category
"good" or "average" for the men as for the females. Thus, in 
*average over all female participants* the number of poor exams is quite correct. For the males,
the number or poor exams is overestimated. For men as well for women on the data
is not reliable on the individual level. 

# 3 Summary
Tu sum up, the coding scheme treats the exams of males and females differently. The
number of average exams is overestimated for females and underestimated for men. 
The number of good exams is underestimated for females but quite accurate for men 
although the number sources from the wrong exams. Finally, the number of poor exams
is quite correct for females although it sources from quite a lot wrong exams while
the number of poor exams for the mean is overestimated. Thus, in descriptive 
statistics females appear to have lower performance as they truly have. For men
the data shows the tendency that they show either a good or a poor performance but
nothing in between.

This example shows how the guidance of the coding scheme can be analysed for
different group of materials. This information can help in developing
coding schemes that are free from bias providing a more reliable and valid
source of data for subsequent analysis and conclusion. 


